Accelerating adaptive ultrasound algorithms using graphics processing units \vspace{0.5cm}

### 2014-04-23 ThesisJP/Introduction/introduction.tex

An interesting observation in \textbf{Paper\,I} can be found in Fig.\,10 \todo{Direct ref}. Here we see how the achieved throughput of the covariance estimation kernel is more than three times higher than the theoretical throughput of a high-end \nom{CPU}{Central processing unit} (The target CPU is from 2010). Note that the target GPU, the Nvidia Quadro 6000, is also from 2010. This clearly shows the benefit of modern GPU computing.

@@

The paper does also propose an interesting method for automatic tuning of rendering parameters. Figure 3d \todo{Direct ref} in \textbf{Paper\,IV} actually shows that an optimal threshold could be found based on global and not only local statistics. Even though the visualization based on global statistics has more errors, with respect to endocardium visibility, the tuning of the opacity function is still fully automatic if an iterative search for the minima in Fig.\,3d in \textbf{Paper\,IV} is performed. A global opacity transfer function will insure a rendering which is realistic, but it will still be dependent on a successful tracking for the auto-adjustment to work.

>>>>

An interesting observation in \textbf{Paper\,I} can be found in Fig.\,\ref{code_assess_flops}. Here we see how the achieved throughput of the covariance estimation kernel is more than three times higher than the theoretical throughput of a high-end \nom{CPU}{Central processing unit} (The target CPU is from 2010). Note that the target GPU, the Nvidia Quadro 6000, is also from 2010. This clearly shows the benefit of modern GPU computing.

@@

The paper does also propose an interesting method for automatic tuning of rendering parameters. Figure \ref{fig:vis}d in \textbf{Paper\,IV} actually shows that an optimal threshold could be found based on global and not only local statistics. Even though the visualization based on global statistics has more errors, with respect to endocardium visibility, the tuning of the opacity function is still fully automatic if an iterative search for the minima in Fig.\,\ref{fig:vis}d is performed. A global opacity transfer function will insure a rendering which is realistic, but it will still be dependent on a successful tracking for the auto-adjustment to work.

### 2014-04-23 ThesisJP/Paper3/moving_obj.tex

The next section gives an introduction to Capon beamforming together with a short introduction to lateral sampling and shift-invariance. In Section \ref{sec:capon_LLSI} we study the scalloping loss yield by the Capon beamformer and its local lateral shift-invariance property.  Different methods for reduced scalloping loss and by that improved local lateral shift-invariance are presented in Section \ref{sec:methods}. Then, in Section \ref{sec:res} we present the results of applying the most promising method on \textit{in vitro} data. Finally we discuss our findings in Section \ref{sec:dis} and give our conclusions in Section \ref{sec:con}.

@@

\begin{figure*}[!t]
\centerline{
\subfloat[]{\includegraphics[width=0.5\linewidth]{\img fig2a.eps}\label{fig:das_shiftinvariant}}
\hfill{}
\subfloat[]{\includegraphics[width=0.5\linewidth]{\img fig2b.eps}\label{fig:das_shiftvariant}}
}
\caption{LSV-plots of delay-and-sum beamforming. If contours are diagonal and the amplitude variation among the peak points (black dots) is small, the imaging system is said to be locally laterally shift-invariant. Dotted vertical lines are plotted where transmit-receive beams are located. a) Sampling density according to (\ref{eq:resolution}). The system is locally laterally shift-invariant. b) Undersampling by a factor of two. The system is no longer locally laterally shift-invariant.}
\label{fig:das}
\end{figure*}

@@

\begin{figure*}[!t]
\centerline{
\subfloat[]{
\includegraphics[width=0.5\linewidth]{\img fig3a.eps}\label{fig:capon1}
}
\subfloat[]{
\includegraphics[width=0.5\linewidth]{\img fig3b.eps}\label{fig:capon2}
}
}
\caption{LSV-plots of Capon beamforming with $L=M/2=48$ and $K=1$, and with sampling corresponding to (\ref{eq:resolution}). a) Capon beamforming with $d=1/100$ as diagonal loading. The system is not locally laterally shift-invariant. b) The diagonal loading factor $d$ is increased to 1. The scalloping loss is reduced, but the system is still not locally laterally shift-invariant.}
\label{fig:capon}
\end{figure*}

@@

\begin{figure*}[!t]
\centerline{
\subfloat[]{
\includegraphics[width=0.5\linewidth]{\img fig4a.eps}\label{fig:capon_oversampling_a}
}
\subfloat[]{
\includegraphics[width=0.5\linewidth]{\img fig4b.eps}\label{fig:capon_oversampling_b}
}
}
\caption{Capon beamforming as described in Fig.\,\ref{fig:capon} with oversampling on transmit. The figure is zoomed compared with Fig.\,\ref{fig:capon} to better see the beam-to-beam variation. a) Four times oversampling ($q=4$). b) Sixteen times oversampling ($q=16$). The maximum scalloping loss is reduced, but it is still too high for the system to be locally laterally shift invariant.}
\label{fig:capon_oversampling}
\end{figure*}

@@

\begin{figure*}[!t]
	\centerline{
		\subfloat[]{
			\includegraphics[width=0.5\linewidth]{\img fig5a.eps}
		}
		\subfloat[]{
			\includegraphics[width=0.5\linewidth]{\img fig5b.eps}
		}
	}
	\caption{Capon beamforming as described in Fig.\,\ref{fig:capon} with oversampling based on phase rotation. The figure is zoomed compared with Fig.\,\ref{fig:capon} to better see the beam-to-beam variation. a) Four times oversampling on receive ($q=4$). b) Sixteen times oversampling on receive ($q=16$).}
	\label{fig:capon_oversampling_pr}
\end{figure*}

\section{In Vitro Phantom Data}\label{sec:res}
In Fig.\,\ref{fig:LSV_invitro} we present the result of applying the Capon beamformer on \textit{in vitro} data with and without oversampling by phase rotation. The data was acquired using a modified Vivid E9 high-end cardiac ultrasound scanner by GE Vingmed Ultrasound, Horten, Norway, equipped with a M5S-D phased transducer operating in harmonic mode at 3.34 MHz. The transmit beam density was set to match (\ref{eq:resolution}) based on this frequency, and 111 tx-beams where therefore distributed on a 65\degree sector. Transmit focus is located at 8 cm range. Note that in harmonic imaging the generated transmit beam at the harmonic frequency will be wider and have lower sidelobes than the beam resulting from transmitting at the harmonic frequency \cite{Fedewa2004}. The lateral sampling is therefore increased compared with our simulations. A tissue mimicking phantom containing wire targets and anechoic cysts was imaged, and the data was acquired while the probe was moved across the phantom manually. Fig.\,\ref{fig:phantom_das_image} shows the first frame out of 14 in the acquired loop processed with DAS beamforming. %Movies of the whole loop processed with DAS and Capon beamforming with and without ten times oversampling by phase rotation can be found in the supplementary material \multimedia{Videos}.

@@

\begin{figure*}[!t]
\centerline{
\subfloat[]{\includegraphics[width=0.5\linewidth]{\img fig7a.png}\label{fig:phantom_capon_image}}
\subfloat[]{\includegraphics[width=0.5\linewidth]{\img fig7b.png}\label{fig:phantom_capon_16PR_image}}
}
\caption{First frame from Fig.\,\ref{fig:phantom_das_image} processed with Capon beamforming. a) Without oversampling \multimedia{Media Movie 2}. b) With sixteen times oversampling using phase rotation \multimedia{Media Movie 3}. Notice how the point scatterers appear brighter in the image with oversampling.}
\label{fig:img_invitro}
\end{figure*}

@@

\section{Discussion}\label{sec:dis}
%Note that in this paper we have investigated a single array configuration and a limited parameter range for the Capon beamformer. The goal was not to make a reference for which oversampling factor to use in different scenarios, but rather do a detailed investigation of the scalloping loss effect. We hope with this investigation that future readers will be able to find out them self what oversampling factor to use for their system.

@@

\section{Conclusion}\label{sec:con}
%In this paper we have investigated the lateral shift-invariance property of the Capon beamformer when imaging moving objects. We have shown that the Capon beamformer is less sensitive to small lateral shifts if oversampling based on phase rotation is applied. This without reducing the acquisition frame rate and with only a minor increase in processing complexity. %In Section \ref{sec:capon_LLSI} it was shown that a Capon beamformer with high resolution parameters is not locally laterally shift-invariant when subjected to motion. The root cause was found to be too coarse lateral sampling. 
%In Section \ref{sec:methods} we presented methods for improved shift-invariance based on oversampling. The most promising method, with respect to improved shift-invariance, reduction in acquisition frame rate and processing complexity combined, was based on phase rotation. 
%In Section \ref{sec:res} we presented results showing that this method also works on \textit{in vitro} data when the ultrasound probe is subjected to motion.

>>>>

The next section gives an introduction to Capon beamforming together with a short introduction to lateral sampling and shift-invariance. In Section \ref{sec:capon_LLSI} we study the scalloping loss yield by the Capon beamformer and its local lateral shift-invariance property.  Different methods for reduced scalloping loss and by that improved local lateral shift-invariance are presented in Section \ref{sec:methods}. Then, in Section \ref{sec:res_paper3} we present the results of applying the most promising method on \textit{in vitro} data. Finally we discuss our findings in Section \ref{sec:dis_paper3} and give our conclusions in Section \ref{sec:con_paper3}.

@@

\begin{figure*}[!t]
\centerline{
\begin{tabular}{c}
\subfloat[]{\includegraphics[width=0.7\linewidth]{\img fig2a.eps}\label{fig:das_shiftinvariant}}\\
\subfloat[]{\includegraphics[width=0.7\linewidth]{\img fig2b.eps}\label{fig:das_shiftvariant}}
\end{tabular}
}
\caption{LSV-plots of delay-and-sum beamforming. If contours are diagonal and the amplitude variation among the peak points (black dots) is small, the imaging system is said to be locally laterally shift-invariant. Dotted vertical lines are plotted where transmit-receive beams are located. a) Sampling density according to (\ref{eq:resolution}). The system is locally laterally shift-invariant. b) Undersampling by a factor of two. The system is no longer locally laterally shift-invariant.}
\label{fig:das}
\end{figure*}

@@

\begin{figure*}[!t]
\centerline{
\begin{tabular}{c}
\subfloat[]{
\includegraphics[width=0.7\linewidth]{\img fig3a.eps}\label{fig:capon1}
}\\
\subfloat[]{
\includegraphics[width=0.7\linewidth]{\img fig3b.eps}\label{fig:capon2}
}
\end{tabular}
}
\caption{LSV-plots of Capon beamforming with $L=M/2=48$ and $K=1$, and with sampling corresponding to (\ref{eq:resolution}). a) Capon beamforming with $d=1/100$ as diagonal loading. The system is not locally laterally shift-invariant. b) The diagonal loading factor $d$ is increased to 1. The scalloping loss is reduced, but the system is still not locally laterally shift-invariant.}
\label{fig:capon}
\end{figure*}

@@

\begin{figure*}[!t]
\centerline{
\begin{tabular}{c}
\subfloat[]{
\includegraphics[width=0.7\linewidth]{\img fig4a.eps}\label{fig:capon_oversampling_a}
}\\
\subfloat[]{
\includegraphics[width=0.7\linewidth]{\img fig4b.eps}\label{fig:capon_oversampling_b}
}
\end{tabular}
}
\caption{Capon beamforming as described in Fig.\,\ref{fig:capon} with oversampling on transmit. The figure is zoomed compared with Fig.\,\ref{fig:capon} to better see the beam-to-beam variation. a) Four times oversampling ($q=4$). b) Sixteen times oversampling ($q=16$). The maximum scalloping loss is reduced, but it is still too high for the system to be locally laterally shift invariant.}
\label{fig:capon_oversampling}
\end{figure*}

@@

\begin{figure*}[!t]
	\centerline{
		\begin{tabular}{c}
		\subfloat[]{
			\includegraphics[width=0.7\linewidth]{\img fig5a.eps}
		}\\
		\subfloat[]{
			\includegraphics[width=0.7\linewidth]{\img fig5b.eps}
		}
		\end{tabular}
	}
	\caption{Capon beamforming as described in Fig.\,\ref{fig:capon} with oversampling based on phase rotation. The figure is zoomed compared with Fig.\,\ref{fig:capon} to better see the beam-to-beam variation. a) Four times oversampling on receive ($q=4$). b) Sixteen times oversampling on receive ($q=16$).}
	\label{fig:capon_oversampling_pr}
\end{figure*}

\section{In Vitro Phantom Data}\label{sec:res_paper3}
In Fig.\,\ref{fig:LSV_invitro} we present the result of applying the Capon beamformer on \textit{in vitro} data with and without oversampling by phase rotation. The data was acquired using a modified Vivid E9 high-end cardiac ultrasound scanner by GE Vingmed Ultrasound, Horten, Norway, equipped with a M5S-D phased transducer operating in harmonic mode at 3.34 MHz. The transmit beam density was set to match (\ref{eq:resolution}) based on this frequency, and 111 tx-beams where therefore distributed on a 65\degree sector. Transmit focus is located at 8 cm range. Note that in harmonic imaging the generated transmit beam at the harmonic frequency will be wider and have lower sidelobes than the beam resulting from transmitting at the harmonic frequency \cite{Fedewa2004}. The lateral sampling is therefore increased compared with our simulations. A tissue mimicking phantom containing wire targets and anechoic cysts was imaged, and the data was acquired while the probe was moved across the phantom manually. Fig.\,\ref{fig:phantom_das_image} shows the first frame out of 14 in the acquired loop processed with DAS beamforming. %Movies of the whole loop processed with DAS and Capon beamforming with and without ten times oversampling by phase rotation can be found in the supplementary material \multimedia{Videos}.

@@

\begin{figure*}[!t]
\centerline{
\begin{tabular}{c}
\subfloat[]{\includegraphics[width=0.7\linewidth]{\img fig7a.png}\label{fig:phantom_capon_image}}\\
\subfloat[]{\includegraphics[width=0.7\linewidth]{\img fig7b.png}\label{fig:phantom_capon_16PR_image}}
\end{tabular}
}
\caption{First frame from Fig.\,\ref{fig:phantom_das_image} processed with Capon beamforming. a) Without oversampling \multimedia{Media Movie 2}. b) With sixteen times oversampling using phase rotation \multimedia{Media Movie 3}. Notice how the point scatterers appear brighter in the image with oversampling.}
\label{fig:img_invitro}
\end{figure*}

@@

\section{Discussion}\label{sec:dis_paper3}
%Note that in this paper we have investigated a single array configuration and a limited parameter range for the Capon beamformer. The goal was not to make a reference for which oversampling factor to use in different scenarios, but rather do a detailed investigation of the scalloping loss effect. We hope with this investigation that future readers will be able to find out them self what oversampling factor to use for their system.

@@

\section{Conclusion}\label{sec:con_paper3}
%In this paper we have investigated the lateral shift-invariance property of the Capon beamformer when imaging moving objects. We have shown that the Capon beamformer is less sensitive to small lateral shifts if oversampling based on phase rotation is applied. This without reducing the acquisition frame rate and with only a minor increase in processing complexity. %In Section \ref{sec:capon_LLSI} it was shown that a Capon beamformer with high resolution parameters is not locally laterally shift-invariant when subjected to motion. The root cause was found to be too coarse lateral sampling. 
%In Section \ref{sec:methods} we presented methods for improved shift-invariance based on oversampling. The most promising method, with respect to improved shift-invariance, reduction in acquisition frame rate and processing complexity combined, was based on phase rotation. 
%In Section \ref{sec:res} we presented results showing that this method also works on \textit{in vitro} data when the ultrasound probe is subjected to motion.
